---
title: <span style="color:blue">Exploratory Data Analysis</span>
author: "Chaden Khalaf(<chaden.khalaf@gmail.com>)"
date: "2024-02-02"
output: html_document
---
```{r echo=FALSE, warning=FALSE}
library(DT)
linelist <- read.csv("linelist.csv") 
linelist %>% 
  datatable(class="cell-border" ,
            caption='Interactive table view of the dataset',
            rowname = F,
            filter = 'top',
            options = list(pageLength = 5, autoWidth = TRUE, scrollX = TRUE))
```{r echo=FALSE, warning=FALSE, fig.cap = "To show the distribution of the Age of the participants of the dataset, a density plot has been used. As it can be seen, major chunk of the sample belongs to the very youth category as a huge proportion lies between the age of 18 and 27 years."}

library(ggplot2)

linelist <- read.csv("linelist.csv") 

ggplot(linelist, aes(x = Age)) + 
  geom_histogram(aes(y = ..density..), colour = "black", fill = "white", bins = 35) +
  geom_density(alpha = 0.5, fill = "blue") + 
  theme_classic() + 
  ylab("Density") +
  ggtitle("Fig1 Age distribution")

```{r echo=FALSE, warning=FALSE, fig.cap = "Interestingly, the dataset contains information about all the obesity levels quite uniformly. Here the most striking thing is that the people belonging to the Normal Weight category only constitutes about 13% of the whole sample. This clearly states that there is an urgent need for implementing strategies to promote healthy lifestyles."}

library(ggplot2)

linelist <- read.csv("linelist.csv") 
ggplot(linelist, aes(x = Obesity_level, fill = Obesity_level)) + geom_bar(stat = "count") + ggtitle("Fig2 Distribution across different obesity levels") + xlab("Obesity level") + ylab("Number of records") + theme_bw() + geom_text(aes(label = ..count..), stat = 'count', vjust = -0.4) + labs(fill = "Obesity level") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))


```{r echo=FALSE, warning=FALSE, fig.cap = "The above plot clearly suggests that most number of males belong to the “Obesity type 2” category, whereas more females belong to the maximum level of Obesity. Interestingly, the difference in numbers is not considerable in the “Normal Weight” category."}

library(ggplot2)

linelist <- read.csv("linelist.csv") 
ggplot(linelist, aes(x = Obesity_level, fill = Gender)) + geom_bar(position = position_dodge(), stat = "count") + theme_bw() + geom_text(aes(label = ..count..), stat = 'count', vjust = -0.4, position = position_dodge(0.9)) + xlab("Obesity level") + ggtitle("Fig3 Gender wise distribution") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.cap = "Most of the people who were either Overweight or Obese had family history with Overweight, as suggested by the Mosaic plot. Now, this is a very crucial information which suggests that there is some relation between the family history and the current physical condition of a person."}
invisible(system("Rscript -e \"install.packages('ggmosaic', quiet = TRUE, repos='https://cloud.r-project.org/')\"", intern = TRUE))
library(ggmosaic)
library(ggplot2)
ggplot(linelist) +
  geom_mosaic(aes(x = product(Family_history_with_overweight, Obesity_level), 
                  fill = Family_history_with_overweight)) +
  xlab("Obesity level") + ylab("Family history with overweight") +
  labs(fill = "Family history with overweight?") + 
  ggtitle("Fig4 Does family history with overweight leads to obesity level?") + theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1))
```{r echo=FALSE, warning=FALSE, message=FALSE, fig.cap = "The above chart might provide an insight on what could be the reason behind the ever-increasing issue. As it can be seen, very few people prefer “Walking” or riding a bike for travelling. Most of the people prefer Public transportation or Automobile modes of transportation."}
ggplot(linelist) +
  geom_bar(aes(y = Transportation_mode,
               fill = Gender),
           position = position_dodge()) + ylab("Mode of Transport") +
  scale_color_manual(values = c("pink", "blue"), aesthetics = "fill") +
  ggtitle("Fig5 Transport preferences by Gender") + theme_mosaic()

